\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-21 paper/paper.tex
Updates to paper.

\section{Introduction}

\begin{figure}
\centering
\includegraphics[scale=0.6]{results2/problem.pdf}
\caption{Rocksdb Write Ahead Logging Impact}
\label{fig:wal_impact}
\end{figure}

\section{Two Level Log}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone'e construction is predicated on the observation that
in the common case protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in persistent directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flash log. We periodically drain entries in batches from the
NVM log to the backing flashlog. An important and beneficial side effect of this
two level approach is that we need only limited amounts of NVM with the bulk of
the log residing on flash - an arrangement that maps well to the relative cost
of the two types of memory. We begin by describing the NVM log and how we
implement the RAFT consensus protocol in the packet switching paradigm around
it. We then describe the flashlog and how we drain entries in batches from the
NVM level to the flash level.

@@

\section{Horizontal Scaling}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated CPU core. Further, we assume a concurrent
key value store where a set of CPU cores independently run operations from the
physical logs on the shared memory key-value store data structure, synchronizing
as necessary.

@@

\section{Evaluation}
We evaluate Cyclone on a 12 node cluster connected via a 10 GigE switch. Three
of the machines are equipped with 1.6TB Intel DC P3600 SSDs and 4*10 GigE
ports. The remaining nine machines do not have SSDs and have only one 10 GigE
port, serving as clients for most of the experiments. As with other
work~\cite{faast}, we use DRAM on the machines to proxy for NVDIMMs where
necessary - the persistent memory needed never exceeds 64 MB regardless of the
size of the key value store or second level log on flash. We divide the
evaluation into three parts. First, we evaluate Cyclone's performance with a
single level log as a pure software packet switch. Next, we evaluate performance
when adding a second level of log on flash. Finally, we evaluate performance
when integrated with Rocksdb as an alternative to Rocksdb's write ahead
log. Unless otherwise mentioned, we use a 60 byte header followed by a payload
for experiments. We log both the header and payload. In all cases the server
echoes the received entry (header and payload) back to the client. We fix the
number of application threads at 32 using at most 8 threads for running RAFT
instances.

>>>>

\section{Introduction}
Persistent key value stores are an increasingly important component of
datacenter scale storage services. Key value stores such as
Rocksdb~\cite{rocksdb}, LevelDB~\cite{leveldb} and FloDB~\cite{flodb} represent
large amounts of effort on both the engineering and deployment fronts. These key
value stores represent sophisticated in-memory data structures built around Log
Structured Merge trees~\cite{lsmtree} and are heavily tuned to extract maximum
performance from present and future generations of flash solid state drives
(SSDs).

These key value stores however have a hitherto ignored an important component -
the write ahead log. A machine or storage failure leading to total loss or
temporary unavailability of data is unacceptable in many internet scale services
where high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write-ahead log that if replicated
and made durable for every appended update provides the necessary high
availability. Unfortunately the write ahead is a performance achilles heel for
these systems - rendering much of the work on improving performance of the LSM
component moot. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked Rocksdb shows the performance of
Rocksdb without the write ahead log. Both persisting the write ahead log without
replicating it - the line marked Rockdb/WAL as well simply replicating the log
three ways using the RAFT~\cite{raft} protocol over TCP/IP without persisting it
are at least an order of magnitude slower (observe the log scale on the
x-axis). It is therefore no suprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store researchers such as the FloDB~\cite{flodb} authors turn off the write
ahead log to be able to showcase benefits of sophisticated extensions to LSM
datastructures.

\begin{figure}
\centering
\includegraphics[scale=0.6]{results2/problem.pdf}
\caption{Rocksdb Write Ahead Logging Impact}
\label{fig:problem}
\end{figure}

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone provides the
order of magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the
server - such as in the form of NVDIMMs~\cite{farm} or directly attached
persistent memory~\cite{pmfs}.

Cyclone makes use of a limited amount of attached NVM using two simple
ideas. First it leverages the observation that with persistence being a first
class property of a subset of directly addressable DRAM, replicating the same
log entry from the leader to follower replicas is no different from multicast
using software packet switching techniques. It directly leverages this
observation to implement a version of RAFT the replicates the NVM log using
classic packet switching techniques such as separation of control and
dataplanes. Second, Cyclone uses a two level log structure where it drains
entries from the NVM log in IO friendly batches to a log place on a flash
drive. This allows Cyclone to use only a limited amount of NVM as current forms
of NVM such as NVDIMMs are more expensive than DRAM. Our experiments use an NVM
footprint of a mere 64 MB to achieve performance comparable to placing the
entire log in NVM. This allows Cyclone to be used in deployments where the key
value store in maintanied entirely in main memory with the log serving as the
source for recovery in the event of failure, rather than a multi-level LSM tree
structure.

We discuss the two level structure focusing on the packet switching components
in Section~\ref{sec:2level}. We then describe how we can horizontally scale this
basic idea across multiple cores by leveraging the commutativity of key value
store APIs across different keys in Section~\ref{sec:horizontal}. We then
demonstrate in the evaluation in Section~\ref{sec:evaluation} that Cyclone can
replicate millions of updates a second using only commodity 10 Gigabit ethernet
thereby closing the performance gap between key value store performance and
write ahead logging performance using only a small amount of directly attached
NVM. A discussion of related work follows before we conclude.

\section{Two Level Log}
\label{sec:2level}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone'e construction is predicated on the observation that
in the common case protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in persistent directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flash log. We periodically drain entries in batches from the
NVM log to the backing flashlog. An important and beneficial side effect of this
two level approach is that we need only limited amounts of NVM with the bulk of
the log residing on flash - an arrangement that maps well to the relative cost
of the two types of memory. We begin by describing the NVM log and how we
implement the RAFT consensus protocol in the packet switching paradigm around
it. We then describe the flashlog and how we drain entries in batches from the
NVM level to the flash level.

@@

\section{Horizontal Scaling}
\label{sec:horizontal}
The basic principle in Cyclone is to turn log replication into a software packet
switching problem. A key ingredient in software packet switching is applying
multiple CPU cores to parallelize the control plane, scaling message processing
throughput. In this section we describe how we can horizontally scale Cyclone's
single \emph{logical} log across multiple \emph{physical} logs. For each
physical log, we run an independent version of RAFT with the corresponding
control plane mapped to a dedicated CPU core. Further, we assume a concurrent
key value store where a set of CPU cores independently run operations from the
physical logs on the shared memory key-value store data structure, synchronizing
as necessary.

@@

\section{Evaluation}
\label{sec:evaluation}
We evaluate Cyclone on a 12 node cluster connected via a 10 GigE switch. Three
of the machines are equipped with 1.6TB Intel DC P3600 SSDs and 4*10 GigE
ports. The remaining nine machines do not have SSDs and have only one 10 GigE
port, serving as clients for most of the experiments. As with other
work~\cite{faast}, we use DRAM on the machines to proxy for NVDIMMs where
necessary - the persistent memory needed never exceeds 64 MB regardless of the
size of the key value store or second level log on flash. We divide the
evaluation into three parts. First, we evaluate Cyclone's performance with a
single level log as a pure software packet switch. Next, we evaluate performance
when adding a second level of log on flash. Finally, we evaluate performance
when integrated with Rocksdb as an alternative to Rocksdb's write ahead
log. Unless otherwise mentioned, we use a 60 byte header followed by a payload
for experiments. We log both the header and payload. In all cases the server
echoes the received entry (header and payload) back to the client. We fix the
number of application threads at 32 using at most 8 threads for running RAFT
instances.
